---
title: "Case Note Analysis: Three-Layer Analytical Framework"
subtitle: "Synthetic Social Services Data Exploration"
author: "Strategic Data Analytics Unit"
date: "`r Sys.Date()`"
format:
  html:
    page-layout: full
    toc: true
    toc-location: right
    toc-depth: 3
    code-fold: show
    code-tools: true
    theme: cosmo
    highlight-style: breeze
    code-line-numbers: true
    df-print: paged
    embed-resources: true
    standalone: true
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
#| code-summary: Reconciles working directory between execution modes
#| echo: false
#| results: hide
#| message: false 
#| warning: false
cat("Working directory: ", getwd())
library(knitr)
opts_knit$set(root.dir='../../') # keep this chunk isolated  
```

```{r}
#| label: set_options
#| code-summary: report-wide formatting options
#| echo: false
#| results: hide
#| message: false 
cat("Working directory: ", getwd()) # turn on to test the location
report_render_start_time <- Sys.time()
# set options shared by all chunks
opts_chunk$set(
  results      = 'show',
  message      = FALSE,
  warning      = FALSE,
  comment      = NA,
  tidy         = FALSE,
  out.width  = "960px", # pixels, this affects only the markdown, not the underlying png file.  The height will be scaled appropriately.
  fig.width    = 12, # inches - wider for multiple panels
  fig.height   = 8, # inches - taller for complex visualizations
  fig.path     = 'figure-png-iso/' # where figures are stored
)
echo_chunks    <- FALSE #Toggle for debugging.
message_chunks <- FALSE #Toggle for debugging.
options(width=120) # number of characters to display in the output (dflt = 80)
ggplot2::theme_set(ggplot2::theme_bw()) # common theme for all graphs
read_chunk("./analysis/eda-2-casenote/eda-2-casenote.R") #This allows knitr to call chunks from R script
```

# Mission

This document explores a **comprehensive three-layer analytical framework** for synthetic social services case notes, progressing from basic demographic profiling through advanced risk assessment to contextual interpretation within reference groups.

**Primary Objectives:**
- **Layer 1**: Establish demographic baseline and univariate distributions  
- **Layer 2**: Create analytical variables and risk flagging systems
- **Layer 3**: Position individual cases within contextual reference frameworks

The analysis supports evidence-based decision making in social services through structured risk assessment and comparative benchmarking.

# Definition of Terms

**Three-Layer Framework**: Progressive analytical approach moving from descriptive to predictive to contextual analysis.

**Risk Flags**: Binary indicators for key intervention areas:
- **Crisis**: Emergency language, urgent intervention needs
- **Housing**: Eviction, homelessness, housing instability  
- **Substance**: Alcohol, drugs, addiction references
- **Mental Health**: Depression, anxiety, suicidal ideation
- **Employment**: Job loss, unemployment concerns
- **Family**: Custody, domestic issues, family separation

**Risk Score**: Composite measure (0-1 scale) representing proportion of risk flags present per case.

**Reference Groups**: Demographic and contextual cohorts used for benchmarking individual cases against relevant baselines.

# Environment

> Reviews the components of the working environment of the report. Non-technical readers are welcomed to skip. Come back if you need to understand the origins of custom functions, scripts, or data objects.

```{r}
#| label: load-packages
#| code-summary: Packages used in the current report
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
```

```{r}
#| label: load-sources
#| code-summary: Collection of custom functions used in current repository  
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
```

```{r}
#| label: declare-globals
#| code-summary: Values used throughout the report.
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
#| eval: true
```

```{r}
#| label: declare-functions
#| code-summary: Custom functions defined for use in this report.
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
#| eval: true
```

# Data

This report operates with **synthetic case notes** generated by the case-note-simulator project, designed to replicate realistic social services documentation while maintaining complete privacy protection.

```{r }
#| label: load-data
#| code-summary: Load synthetic case notes dataset from project output
#| warning: false
#| code-fold: true
#| cache: true
```

```{r}
#| label: inspect-data
#| code-summary: High-level overview of synthetic case notes dataset
#| code-fold: false
#| eval: true
#| cache: true
```

# Layer 1: Input Data Description & Basic Demographics

> **Layer 1 Mission**: Establish comprehensive understanding of input data through systematic univariate analysis of all non-text variables, creating the foundation for subsequent analytical transformations.

## Data Preparation & Overview

```{r}
#| label: layer-1-data-preparation
#| code-summary: Prepare dataset with standardized variables and age groups
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Q1.1: Age Distribution Analysis

> Q1.1: What is the age distribution of our synthetic case note population?

```{r}
#| label: layer-1-univariate-age
#| code-summary: Age distribution analysis with summary statistics and visualization
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Q1.2: Gender Distribution Analysis

> Q1.2: How is gender distributed across the population?

```{r}
#| label: layer-1-univariate-gender
#| code-summary: Gender distribution analysis with proportions
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Q1.3: Geographic Distribution Analysis

> Q1.3: What is the geographic spread of cases across locations?

```{r}
#| label: layer-1-univariate-location
#| code-summary: Location distribution analysis
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Q1.4: Case Note Characteristics

> Q1.4: What are the length and complexity characteristics of case notes?

```{r}
#| label: layer-1-univariate-notes
#| code-summary: Case note length and word count analysis
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Q1.5: Demographic Relationships

> Q1.5: How do demographic variables relate to each other?

```{r}
#| label: layer-1-bivariate-demographics
#| code-summary: Cross-tabulations and relationships between demographic variables
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Layer 1 Summary

```{r}
#| label: layer-1-summary
#| code-summary: Layer 1 completion summary and data export
#| echo: false
#| results: show
#| message: false
#| warning: false
#| code-fold: false
#| cache: true
```

# Layer 2: Adding Analytical Variables & Transformations

> **Layer 2 Mission**: Transform the input data by adding analytical variables, risk flags, and composite measures that enable advanced case assessment and risk stratification.

## Q2.1: Risk Flag Creation

> Q2.1: What risk patterns can we identify through systematic text analysis of case notes?

```{r}
#| label: layer-2-risk-flags
#| code-summary: Create risk flags and analytical variables from case note content
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Q2.2: Risk Flag Analysis

> Q2.2: What is the prevalence of different risk types across the population and demographic groups?

```{r}
#| label: layer-2-risk-flag-analysis
#| code-summary: Analyze prevalence of risk flags across population and demographic segments
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Q2.3: Risk Flag Visualizations

> Q2.3: How can we visualize risk patterns across demographic groups?

```{r}
#| label: layer-2-visualizations
#| code-summary: Create visualizations for risk flag patterns and distributions
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Q2.4: Composite Risk Scoring

> Q2.4: How do individual cases score on our composite risk measures?

```{r}
#| label: layer-2-risk-score-analysis
#| code-summary: Analyze composite risk score distribution and demographic patterns
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Q2.5: Note Complexity & Writing Patterns

> Q2.5: What writing style patterns and note complexity indicators can we identify?

```{r}
#| label: layer-2-note-complexity
#| code-summary: Analyze note writing patterns and complexity indicators
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Layer 2 Summary

```{r}
#| label: layer-2-summary
#| code-summary: Layer 2 completion summary with key analytical variables added
#| echo: false
#| results: show
#| message: false
#| warning: false
#| code-fold: false
#| cache: true
```

# Layer 3: Contextual Interpretation & Reference Groups

> **Layer 3 Mission**: Position individual cases within meaningful contextual frameworks by establishing population baselines and enabling comparative analysis across demographic and geographic reference groups.

## Q3.1: Population Baseline Establishment

> Q3.1: What are the population-wide baseline rates for all risk indicators?

```{r}
#| label: layer-3-baseline-establishment
#| code-summary: Establish population baselines for contextual comparison
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Q3.2: Age Group Context Analysis

> Q3.2: How do different age groups compare to the population baseline for risk indicators?

```{r}
#| label: layer-3-age-group-context
#| code-summary: Compare each age group to overall baseline risk patterns
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Q3.3: Geographic Context Analysis

> Q3.3: What geographic variations exist in risk patterns across locations?

```{r}
#| label: layer-3-geographic-context
#| code-summary: Analyze geographic variation in risk patterns
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Q3.4: Gender Context Analysis

> Q3.4: How do risk patterns differ by gender relative to population baseline?

```{r}
#| label: layer-3-gender-context
#| code-summary: Analyze gender-based risk profile differences
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Q3.5: Contextual Visualizations

> Q3.5: How can we visualize contextual patterns and baseline comparisons?

```{r}
#| label: layer-3-visualizations
#| code-summary: Create baseline comparison visualizations and heatmaps
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Q3.6: Risk Stratification Framework

> Q3.6: How can we create a systematic risk stratification system for operational use?

```{r}
#| label: layer-3-risk-stratification
#| code-summary: Create risk tier classification and stratification framework
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Q3.7: Individual Case Positioning

> Q3.7: How do high-risk cases position within their reference groups?

```{r}
#| label: layer-3-individual-positioning
#| code-summary: Position individual cases within reference groups
#| echo: true  
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Layer 3 Summary

```{r}
#| label: layer-3-summary
#| code-summary: Layer 3 completion summary with contextual framework established
#| echo: false
#| results: show
#| message: false
#| warning: false
#| code-fold: false
#| cache: true
```

# Comprehensive Analysis Summary

## Final Results Overview

```{r}
#| label: comprehensive-summary
#| code-summary: Comprehensive summary of all three layers with key metrics
#| echo: false
#| results: show
#| message: false
#| warning: false
#| code-fold: false
#| cache: true
```

## Export for Advanced Processing

This analysis prepares the enhanced dataset for subsequent Python NLP processing, enabling advanced sentiment analysis, topic modeling, and machine learning applications.

```{r}
#| label: export-for-python-nlp
#| code-summary: Prepare enhanced dataset for Python NLP processing
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

## Analysis Recommendations & Next Steps

```{r}
#| label: analysis-recommendations
#| code-summary: Generate specific recommendations based on analysis findings
#| echo: false
#| results: show
#| message: false
#| warning: false
#| code-fold: false
#| cache: true
```

## Files Created & Interactive Functions

```{r}
#| label: files-created-summary
#| code-summary: Summary of all files and visualizations created during analysis
#| echo: false
#| results: show
#| message: false
#| warning: false
#| code-fold: false
#| cache: true
```

## Interactive Exploration Tools

The analysis creates several interactive functions for ongoing data exploration:

```{r}
#| label: interactive-functions
#| code-summary: Helper functions for interactive exploration of results
#| echo: true
#| results: show
#| message: false
#| warning: false
#| code-fold: true
#| cache: true
```

# Methodological Framework

## Three-Layer Analytical Approach

This analysis demonstrates a **progressive analytical methodology** that moves systematically from descriptive through analytical to contextual interpretation:

### **Layer 1: Foundation**
- **Purpose**: Establish comprehensive understanding of input data
- **Method**: Systematic univariate analysis of all variables
- **Output**: Demographic baselines and distributional understanding
- **Value**: Ensures analytical decisions are grounded in data reality

### **Layer 2: Transformation**  
- **Purpose**: Create analytical variables enabling advanced assessment
- **Method**: Text analysis, risk flagging, composite measure creation
- **Output**: Enhanced dataset with risk indicators and behavioral flags
- **Value**: Transforms raw data into actionable analytical framework

### **Layer 3: Contextualization**
- **Purpose**: Position findings within meaningful reference frameworks
- **Method**: Baseline comparison, reference group analysis, individual positioning
- **Output**: Contextual interpretation enabling comparative benchmarking
- **Value**: Enables evidence-based decision making through relative assessment

## Technical Implementation

**R Environment**: Comprehensive statistical analysis and visualization using tidyverse ecosystem, with structured chunk-based organization supporting reproducible research.

**Python Integration**: Analysis prepares enhanced datasets for advanced NLP processing including sentiment analysis, topic modeling, and machine learning applications.

**Quarto Reporting**: Dynamic document generation enabling both technical documentation and stakeholder communication through flexible output formatting.

## Data Quality & Limitations

**Synthetic Data Benefits**:
- Complete privacy protection while maintaining analytical utility
- Controlled distributions enabling systematic validation
- Realistic variation reflecting authentic caseworker patterns

**Analytical Considerations**:
- Basic keyword detection provides foundation for advanced NLP enhancement
- Risk scoring methodology balances simplicity with meaningful differentiation  
- Contextual interpretation depends on appropriate reference group selection

## Operational Applications

This framework supports:
- **Strategic Planning**: Population-level risk assessment and resource allocation
- **Case Management**: Individual risk flagging and intervention prioritization
- **Quality Assurance**: Systematic assessment of case documentation patterns
- **Policy Development**: Evidence-based service delivery optimization

---

**Report Generation Complete**: `r Sys.time()`\
**Processing Time**: `r round(difftime(Sys.time(), report_render_start_time, units = "mins"), 1)` minutes\
**Framework**: Three-layer analytical approach with comprehensive risk assessment capabilities\

This three-layer framework provides a replicable methodology for systematic analysis of social services case documentation, supporting both research objectives and operational decision-making requirements.