---
title: "Case Note Analysis: Three-Layer Analytical Framework"
subtitle: "Synthetic Social Services Data Exploration"
author: "Strategic Data Analytics Unit"
date: "`r Sys.Date()`"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    code-fold: show
    code-tools: true
    df-print: paged
    embed-resources: true
    standalone: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::read_chunk("eda-2-casenote.R")
```

# Executive Summary

This report presents a three-layer analytical framework for examining synthetic case notes from social services clients. The analysis progresses from population-level demographic patterns through individual risk assessment to contextual interpretation within reference groups.

**Key Findings Preview:**
- Population characteristics and risk stratification patterns
- Individual case note risk flagging capabilities  
- Contextual benchmarking against relevant comparison groups

# Analytical Framework

## Layer 1: Demographic Analysis
Population-level profiling focusing on risk stratification across client characteristics.

## Layer 2: Individual Case Analysis  
Note-level examination with NLP-enhanced risk flagging for substance use, housing instability, and mental health indicators.

## Layer 3: Contextual Interpretation
Comparative analysis positioning individual cases within meaningful reference groups.

# Data Overview

```{r environment}
```

```{r load-data}
```

```{r inspect-data}
```

# Layer 1: Population Demographics

```{r layer-1-demographics}
```

The demographic analysis reveals the composition and risk profile of our synthetic client population. Key stratification factors include age distribution, geographic location, and baseline complexity levels.

**Risk Stratification Insights:**
- Distribution across complexity levels provides framework for resource allocation
- Geographic patterns may indicate service accessibility challenges
- Age demographics inform service delivery approaches

# Layer 2: Individual Case Analysis

```{r layer-2-individual}
```

Individual case note analysis combines basic text characteristics with NLP-ready preparation for advanced risk flagging.

**Risk Flag Development:**
- Crisis indicators: Emergency language and urgent intervention needs
- Housing stability: Eviction notices, homelessness references  
- Substance use: Direct mentions and related terminology
- Mental health: Depression, anxiety, suicidal ideation indicators

**Note:** Advanced sentiment analysis and nuanced risk scoring will be implemented via Python NLP models in subsequent processing stages.

# Layer 3: Contextual Interpretation  

```{r layer-3-contextual}
```

Contextual analysis positions individual cases within appropriate reference frameworks, enabling meaningful comparison and benchmarking.

**Reference Group Framework:**
- **Overall Population**: Baseline rates for all risk indicators
- **Complexity Cohorts**: Comparison within similar case complexity levels
- **Geographic Context**: Regional patterns and service environment factors

# Visualizations

## Demographic Patterns

```{r visualization}
```

## Risk Flag Analysis

The risk flagging system provides structured approach to identifying key intervention needs across the client population.

**Interpretation Guidelines:**
- Higher complexity levels should show elevated risk flag prevalence
- Geographic differences may indicate regional service gaps
- Individual cases can be benchmarked against appropriate reference groups

# Methodological Notes

## Data Quality Considerations
- Synthetic data maintains realistic variation while ensuring complete privacy protection
- Generated notes reflect authentic caseworker writing patterns and terminology
- Controlled complexity distributions enable systematic analysis validation

## Analytical Limitations
- Basic keyword detection pending enhancement with sophisticated NLP models
- Sentiment analysis capabilities require Python integration
- Contextual interpretation depends on appropriate reference group selection

## Next Steps

1. **NLP Enhancement**: Integration with Python models for advanced text analysis
2. **Risk Scoring**: Development of composite risk indicators
3. **Longitudinal Analysis**: Temporal patterns in case progression
4. **Intervention Mapping**: Connection between risk profiles and service recommendations

# Technical Implementation

This analysis demonstrates the integration of:
- R-based statistical analysis and visualization
- Structured approach to multi-layer analytical frameworks  
- Preparation for Python NLP model integration
- Reproducible reporting via Quarto

The framework supports both exploratory research needs and operational decision-making requirements for social services analytics.